---
title: "POWER & SAP"
author: "SP"
date: "2023-03-24"
output:
  html_document:
    df_print: paged
  word_document:
    fig_width: 10
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# This ia a project that I worked on at the Institute of Infectious Diseases Clinic, Pune. The objective of the study was to assess the trends in bodyweight after switching from drug A to drug B/drug b (lower dose of drug B)/continue on drug A. Since the study involved randomizing patients to these three arms, calculating sample size to enable comparisons between three groups was needed. The `ThreeArmedTrials` package is made for this explicit purpose. Please see the R code below to get a rough idea.

## POWER & SAMPLE SIZE

```{r include=FALSE}
# This will load the `ThreeArmedTrials` package
library(ThreeArmedTrials)
```

### OPTIMAL SAMPLE SIZE ALLOCATION CALCULATION

```{r}
# Input effect sizes for drug B (so called "experiment"), drug b (so called "reference"), and drug A arms (so called "placebo"); set a margin (delta), and choose what type of distribution and link function you want; store the optimal allocation it in an object for easier retrieval
w <- ThreeArmedTrials::opt_alloc_RET(experiment = 0.4, 
                                     reference = 0.2, 
                                     placebo = 0.05,
                                     Delta = 0.7,
                                     distribution = "binary", 
                                     h = function(x){-log(x/(1-x))})
print(w)
```

# The optimal sample size allocation for a Wald-type test for a three-armed superiority trial, where the probability rate of patients on drug B who will lose ≥5% weight is estimated to be 40%, for drug b is estimated to be 20%, and for the reference group continuing on drug A is estimated to be 5%, with a superiority margin of a log odds 0.7 (odds ratio of 2) and a binary distribution following a logit link function, is estimated to be 39.5%, 33.9%, and 26.6% approximately for the drug B, drug B, and drug A group respectively.

### SAMPLE SIZE AND POWER RELATED CALCULATIONS

```{r}
# With optimal allocation as calculated in the previous cell
ThreeArmedTrials::power_RET(experiment = 0.4, 
                            reference = 0.2, 
                            placebo = 0.05,
                            sig_level = 0.05, 
                            power = 0.8,
                            Delta = 0.7, 
                            allocation = w, 
                            distribution = "binary", 
                            h = function(x){-log(x/(1-x))}, 
                            h_inv = function(x){exp(-x)/(1+exp(-x))},
                            var_estimation = "RML")
```

# The sample size for a Wald-type test for superiority of transitioning to drug B versus drug b with respect to maintenance on drug A assuming a probability rate for outcomes of 40% in the drug B group, 20% in the drug b group, and 5% in the drug A group, an alpha error of 5%, power at 80%, a superiority margin log odds of 0.7 (odds ratio of 2) is calculated to be 29, 25, and 19 in the drug B, drug b, and drug A arms respectively.

```{r}
# With 1:1:1 allocation (not the optimal allocation)
ThreeArmedTrials::power_RET(experiment = 0.4, 
                            reference = 0.2, 
                            placebo = 0.05,
                            sig_level = 0.05, 
                            power = 0.8,
                            Delta = 0.7, 
                            allocation = c(1, 1, 1)/3, 
                            distribution = "binary", 
                            h = function(x){-log(x/(1-x))}, 
                            h_inv = function(x){exp(-x)/(1+exp(-x))},
                            var_estimation = "RML")
```

# The goal is to see whether transitioning to drug B is clearly superior to transitioning to drug b from drug A with respect to maintaining on drug A in losing clinically significant bodyweight (≥5%) at 48 weeks in patients transitioning from drug A Following this, the primary hypothesis that arises is:
# Whether transitioning to drug B from drug A is superior to transitioning to drug b with respect to maintenance on drug A in losing clinically significant bodyweight (≥5%) at 48 weeks.

# Secondary hypotheses include:
# 1.  Whether transitioning to drug B from drug A is superior to maintenance on drug A in losing clinically significant bodyweight (≥5%) at 48 weeks.
# 2.  Whether transitioning to drug b from drug A is superior to maintenance on drug A in losing clinically significant bodyweight (≥5%) at 48 weeks.
# 3.  Whether transitioning to drug B from drug A is superior to transitioning to drug b from drug A with respect to maintenance on drug A in losing clinically significant bodyweight (≥5%) at 48 weeks after stratifying for type 2 diabetes mellitus.

# Based on clinical judgement, we anticipate that there is 40% probability that patients on drug B will lose clinically significant bodyweight (≥5%) after transitioning from drug A, 20% probability that patients on drug b will lose clinically significant bodyweight (≥5%) after transitioning from drug A , and 5% probability that patients maintained on drug A will lose clinically significant bodyweight (≥5%).

# Sample size is calculated for a superiority three-arm trial with binary endpoints using the logit link function for a Wald-type test using the restricted variance estimation (RML). The allocation ratio is 1:1:1. We decided a log odds superiority margin (delta) of 0.7 (corresponding to an odds ratio of 2) to be clinically significant to investigate the superiority of drug B over drug b The required sample size for achieving 80% power at a significance level of 5% is 27 patients in each group (N = 81). Assuming a dropout rate of approximately 20% over a 48-week period in an urban population, we plan to recruit 35 patients in each group (N = 105).

### TESTING THE RETENTION OF EFFECT HYPOTHESIS

```{r include=FALSE}
# Please note that the datafile contained contains fictional simulated data and NOT the real data of patients
df1 <- read.csv("/your/data/for/testing/the/retention/of/effect/hypothesis/RoE data.csv")
# Define variable types
df1$Experimental <- as.numeric(df1$Outcome)
df1$Group <- as.factor(df1$Group)
```

```{r include=FALSE}
# To change the reference group as needed
df1$Group <- relevel(df1$Group, ref = "Drug b")
```

```{r include=FALSE}
# To change the reference group as needed
df1$Group <- relevel(df1$Group, ref = "Drug A")
```

```{r}
# Test the retention of effect hypothesis
ThreeArmedTrials::test_RET(xExp = df1$Outcome[df1$Group == "Drug B"],
                           xRef = df1$Outcome[df1$Group == "Drug b"],
                           xPla = df1$Outcome[df1$Group == "Drug A"], 
                           Delta = 0.7, 
                           var_estimation = "RML",
                           distribution = "binary",
                           h = function(x){-log(x/(1-x))},
                           h_inv = function(x){exp(-x)/(1+exp(-x))})
```

```{r}
# Perform a manual binary logistic regression
fit <- glm(formula = Outcome ~ Group, family = binomial, data = df1)
summary(fit)
coef(fit)
confint(fit)
```

# We tested the retention of effect hypothesis on a simulated dataset with 1:1:1 allocation and 35 participants in each group. Fourteen patients on drug B, seven patients on drug b, and two patients on drug A lost clinically significant bodyweight (≥5%). A Wald-type test yielded a statistically significant p-value (p = 2.494e-03) between the drug B and drug b groups with respect to drug A indicating that there is a significant difference between the groups.

```{r include=FALSE}
# Plot a figure to illustrate various scenarios regarding confidence interval crossing of the null point and margins well ahead in the analysis plan itself before study commencement
df2 <- read.csv("your/hypothetical/data/for/the/point/estimate/and/confidence/limits/CI figure.csv")
```

```{r include=FALSE}
library(ggplot2)
```

```{r}
# Set the defined margin (0.7) to illustrate various possible scenarios
ggplot2::ggplot(data = df2) + 
         geom_errorbar(aes(xmin = LCL, xmax = UCL, y = ID)) + 
         geom_vline(xintercept = c(0, 0.7), lty = c(1, 4)) + 
         scale_x_continuous(breaks = seq(-0.2, 3.2, 0.2)) +
         theme_minimal() +
         xlab("Log Odds")
```

# In the figure, each point estimate with its CI can be interpreted as follows:
# 1.  Shows the lower confidence limit for the log odds which is far away from the superiority margin (delta) at 0.7 - superiority shown.
# 2.  Shows the lower confidence limit for the log odds which is close, but not crossing the superiority margin (delta) at 0.7 - superiority shown.
# 3.  Shows the confidence interval crossing the superiority margin (delta) at 0.7 - superiority not shown.
# 4.  Shows the confidence interval with the upper confidence limit lower than the superiority margin (delta) at 0.7 - superiority not shown.
# 5.  Shows the confidence interval with the upper confidence limit lower than the superiority margin (delta) at 0.7 - superiority not shown.
# 6.  Shows the confidence interval with the upper confidence limit lower than the superiority margin (delta) at 0.7 - superiority not shown.

## STATISTICAL ANALYSIS PLAN

# The intention-to-treat protocol will be followed to maintain randomization and sample size. The primary endpoint will a comparison of the drug B arm to the drug b arm with respect to the drug A arm in losing clinically significant bodyweight (≥5%). Since the goal is to see whether drug B is superior to drug b in losing clinically significant bodyweight (≥5%) at 48 weeks, the primary hypothesis will be whether transitioning to drug B from drug A is superior to transitioning to drug b from drug A with respect to maintenance on drug A We anticipate the drug B arm to show the highest reduction in bodyweight, the drug A arm to show the least reduction in bodyweight, and the drug b ar to show an intermediate reduction in bodyweight at 48 weeks.

# For the primary outcome, a logistic regression model will be used to compare the odds of achieving clinically significant bodyweight loss (≥5%) with the drug A arm serving as the reference class. The model will be adjusted to the following confounding variables: presence of diabetes mellitus, age, sex, duration on drug A, baseline BMI, and baseline CD4 levels. Linear mixed effects (LME) models with each participant as a random intercept will be used to fit and predict CD4 counts, weight, BMI, wait-hip ratio, blood sugar levels, HbA1c levels, lipid parameters, and estimated GFR over time. Logistic regression models adjusted for confounders will be used to estimate the odds of events with their risk factors. Kaplan-Meier analyses will be performed to estimate median times to events. The log-rank test will be used to compare Kaplan-Meier estimates. Cox regression models adjusted for confounders will be used to get hazard estimates for the occurrence of events over time. The proportional hazards assumption will be tested using Schoenfeld residuals. Chi-square or Fisher's exact tests will be used to compare categorical variables as appropriate. The Cochran-Mantel-Haenszel test will be used for simple stratified analysis. Student's t-tests or analysis of variance techniques (ANOVA) will be used to compare continuous variables for two or more groups respectively. Patients who had achieved events of interest at baseline will be censored from the concerned analyses.

# Subgroup analyses will be performed to assess the robustness and validity of the results, and whether associations are stronger or weaker in different subgroups. Similar statistical methods will be used in subgroup analyses to those used to assess primary and secondary outcomes. Subgroups will include patients with baseline BMI ≥25 kg/m^2 and those with <25 kg/m^2, males and females, family history of obesity and those without, CyP2B polymorphisms and those without, and those with baseline CD4 ≤350 cells/cumm and those >350 cells/cumm.

# For the two co-primary outcomes, p-values less than the alpha significance level of 0.025 (0.05/2) will be considered to be statistically significant. For the rest of the secondary outcomes, p-values less than 0.001 (0.05/50) will be considered to be statistically significant; exact p-values, effect sizes and point estimates with their 95% confidence intervals will be reported.
